Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-11 doc/manuscript/ESSDmanuscript.Rmd
working on intro

To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.
The EFDB is intended as a recognized library, and can be used both for efforts to tally a nation's intended or accomplished greenhouse gas reductions, or as a basis of comparison for external parties to evaluate these inventories. 
The EFDB encourages researchers to submit estimates of emission factors or other related parameters from peer-reviewed journal papers or other accepted sources for inclusion in the database. 
In the case of forests, emission factors refer to ... <!--C stocks and increments, and check other GHGs-->

**The Global Forest Carbon Database, ForC, is the largest collection of published estimates of forest carbon stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021].**
*(add stats/ details, maybe record of how ForC has grown over time)*
As such, ForC is positioned to improve forest C accounting through the transfer of data to EFDB.
The purpose of this publication is to document that process and provide recommendations for future improvements.

**Here, we**
(1) review IPCC definitions of relevant carbon stocks and increments
(2) describe mapping of ForC to IPCC's EFDB,
(3) describe updates to ForC  (ForC v4.0),
(4) summarize the data in ForC that's relevant to EFDB, identifying gaps, and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

>>>>

To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; [https://www.ipcc-nggip.iges.or.jp/EFDB/main.php](https://www.ipcc-nggip.iges.or.jp/EFDB/main.php)), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.
The EFDB can be used both for efforts to tally a nation's intended or accomplished greenhouse gas reductions, or as a basis of comparison for external parties to evaluate these inventories. 
The EFDB encourages researchers to submit estimates of emission factors or other related parameters from peer-reviewed journal papers or other accepted sources for inclusion in the database. 
In the case of forests, emission factors include carbon stocks and increments within various pools (Table&nbsp;\@ref(table_variables)).

The Global Forest Carbon Database, ForC ([https://forc-db.github.io/](https://forc-db.github.io/)), is the largest collection of published estimates of forest carbon stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021].
For C currently contains **##** records from **##** plots in **##** distinct geographical areas, along with records of stand age and disturbance history.
As such, ForC is positioned to improve forest C accounting through the transfer of data to EFDB.
The purpose of this publication is to document that process and provide recommendations for future improvements.

**Here, we**
(1) review IPCC definitions of relevant carbon stocks and increments;
(2) describe mapping of ForC to IPCC's EFDB;
(3) describe updates to ForC  (ForC v4.0), most of which were implemented to facilitate data transfer to EFDB;
(4) summarize the data in ForC that's relevant to EFDB and records that have been transferred to date; and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

### 2022-04-11 doc/manuscript/ESSDmanuscript.tex
working on intro

To improve the data accessible for C accounting, the IPCC created the
Emission Factor Database (EFDB;
https://www.ipcc-nggip.iges.or.jp/EFDB/main.php), which is intended as a
recognized library of emission factors and other parameters that can be
used for estimating greenhouse gas emissions and removals. The EFDB is
intended as a recognized library, and can be used both for efforts to
tally a nation's intended or accomplished greenhouse gas reductions, or
as a basis of comparison for external parties to evaluate these
inventories. The EFDB encourages researchers to submit estimates of
emission factors or other related parameters from peer-reviewed journal
papers or other accepted sources for inclusion in the database. In the
case of forests, emission factors refer to \ldots{}

\textbf{The Global Forest Carbon Database, ForC, is the largest
collection of published estimates of forest carbon stocks, increments,
and annual fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}.}
\emph{(add stats/ details, maybe record of how ForC has grown over
time)} As such, ForC is positioned to improve forest C accounting
through the transfer of data to EFDB. The purpose of this publication is
to document that process and provide recommendations for future
improvements.

\textbf{Here, we} (1) review IPCC definitions of relevant carbon stocks
and increments (2) describe mapping of ForC to IPCC's EFDB, (3) describe
updates to ForC (ForC v4.0), (4) summarize the data in ForC that's
relevant to EFDB, identifying gaps, and (5) provide recommendations for
improving data collection, analysis, database, and accounting.

>>>>

To improve the data accessible for C accounting, the IPCC created the
Emission Factor Database (EFDB;
\url{https://www.ipcc-nggip.iges.or.jp/EFDB/main.php}), which is
intended as a recognized library of emission factors and other
parameters that can be used for estimating greenhouse gas emissions and
removals. The EFDB can be used both for efforts to tally a nation's
intended or accomplished greenhouse gas reductions, or as a basis of
comparison for external parties to evaluate these inventories. The EFDB
encourages researchers to submit estimates of emission factors or other
related parameters from peer-reviewed journal papers or other accepted
sources for inclusion in the database. In the case of forests, emission
factors include carbon stocks and increments within various pools
(Table~@ref(table\_variables)).

The Global Forest Carbon Database, ForC
(\url{https://forc-db.github.io/}), is the largest collection of
published estimates of forest carbon stocks, increments, and annual
fluxes
\citep{anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}. For
C currently contains \textbf{\#\#} records from \textbf{\#\#} plots in
\textbf{\#\#} distinct geographical areas, along with records of stand
age and disturbance history. As such, ForC is positioned to improve
forest C accounting through the transfer of data to EFDB. The purpose of
this publication is to document that process and provide recommendations
for future improvements.

\textbf{Here, we} (1) review IPCC definitions of relevant carbon stocks
and increments; (2) describe mapping of ForC to IPCC's EFDB; (3)
describe updates to ForC (ForC v4.0), most of which were implemented to
facilitate data transfer to EFDB; (4) summarize the data in ForC that's
relevant to EFDB and records that have been transferred to date; and (5)
provide recommendations for improving data collection, analysis,
database, and accounting.
